---
title: "Homework 8"
author: "KS"
date: "2023-03-22"
output: html_document
---
Load in `library(tidyverse)` and `library(ggplot2)`, hidden below.

```{r include=FALSE}
library(tidyverse)
library(ggplot2)
```

## Paper

1. Little, S. & Moore, E.B. 1950 "Effect of Prescribed Burns and Shelterwood Cutting on Reproduction of Shortleaf and Pitch Pine". Northeastern Forest Experiment Station, USFS, Upper Darby, PA. 

## Hypothesis

Number of repeat burns increases the number of pine seedlings present.

## Study design, means, and standard deviation

2. 14 1-acre plots; 2 plots left unburned.  
Other 12, RxFire winter burn applied 1, 2, 3, 4, 5, & 6 times (1940-1946).  
On half of plots in 1939, overstory oaks thinned via sheterwood. 
On all plots, pine at 15 seed trees/acre.  
Remaining oak overstory cut on half of each plot in winter 1946/47. Other half cut winter 1948/49.  

As assessed in 1949,  
-     unburned plots with 580 seedlings/acre  
-     1 burn : 800 seedlings/acre  
-     2 burn : 1,000 seedlings/acre  
-     3 burn : 1,925 seedlings/acre  
-     4 burn : 1,925 seedlings/acre  
-     5 burn : 2,650 seedlings/acre  
-     6 burn : 2,650 seedlings/acre  
  
  Assuming standard deviation is 100 seedlings groups 0-2; 200 seedlings 3-4; 400 seedlings 5-6

## Making random data based on study data

  3. Creating uniform distribution for each treatment group.
```{r}
noBurn <- rnorm(n=2, mean = 580, sd = 100)
oneBurn <- rnorm(n=2, mean = 800, sd = 100)
twoBurn <- rnorm(n=2, mean = 1000, sd = 100)
threeBurn <- rnorm(n=2, mean = 1925, sd = 200)
fourBurn <- rnorm(n=2, mean = 1925, sd = 200)
fiveBurn <- rnorm(n=2, mean = 2650, sd = 400)
sixBurn <- rnorm(n=2, mean = 2650, sd = 400)

pineBurn <- data.frame(noBurn,oneBurn,twoBurn,threeBurn,fourBurn,fiveBurn,sixBurn)

pineBurn <- pineBurn %>% 
  pivot_longer(cols = noBurn:sixBurn,
               names_to = "treatment",
               values_to = "seedling_counts1")
  
pineBurn1 <- pineBurn[with(pineBurn, order(treatment)),]
```
  
4. Create a ANOVA where we compare means

```{r}
A1 <- aov(pineBurn1$seedling_counts1~pineBurn$treatment)
summary(A1)
```
```{r}
ggplot(pineBurn1, aes(treatment, seedling_counts1)) +
  geom_boxplot()
```
This does not yeild a significant p-value

## Changing the means

6. Changing means for a statistically significant result

```{r}
noBurn4 <- rnorm(n=2, mean = 570, sd = 100)
oneBurn4 <- rnorm(n=2, mean = 790, sd = 100)
twoBurn4 <- rnorm(n=2, mean = 990, sd = 100)
threeBurn4 <- rnorm(n=2, mean = 1975, sd = 200)
fourBurn4 <- rnorm(n=2, mean = 1975, sd = 200)
fiveBurn4 <- rnorm(n=2, mean = 2725, sd = 400)
sixBurn4 <- rnorm(n=2, mean = 2725, sd = 400)

pineBurn4 <- data.frame(noBurn4,oneBurn4,twoBurn4,threeBurn4,fourBurn4,fiveBurn4,sixBurn4)

pineBurn4 <- pineBurn4 %>% 
  pivot_longer(cols = noBurn4:sixBurn4,
               names_to = "treatment",
               values_to = "seedling_counts1")
  
pineBurn5 <- pineBurn4[with(pineBurn4, order(treatment)),]
```
  
I decreased the first 3 treatment groups by 10 and increased the final 4 treatment groups by 50 and the p value became significant. I started with a higher value, but it did not take much change to yeild a significant p-value. 

```{r}
A3 <- aov(pineBurn4$seedling_counts1~pineBurn4$treatment)
summary(A3)
```

```{r}
ggplot(pineBurn4, aes(treatment, seedling_counts1)) +
  geom_boxplot()
```



## Increasing the sample size

7. Increasing the sample size for a statistically significant result

```{r}
noBurn2 <- rnorm(n=3, mean = 580, sd = 100)
oneBurn2 <- rnorm(n=3, mean = 800, sd = 100)
twoBurn2 <- rnorm(n=3, mean = 1000, sd = 100)
threeBurn2 <- rnorm(n=3, mean = 1925, sd = 200)
fourBurn2 <- rnorm(n=3, mean = 1925, sd = 200)
fiveBurn2 <- rnorm(n=3, mean = 2650, sd = 400)
sixBurn2 <- rnorm(n=3, mean = 2650, sd = 400)

pineBurn2 <- data.frame(noBurn2,oneBurn2,twoBurn2,threeBurn2,fourBurn2,fiveBurn2,sixBurn2)

pineBurn2 <- pineBurn2 %>% 
  pivot_longer(cols = noBurn2:sixBurn2,
               names_to = "treatment",
               values_to = "seedling_counts1")
  
pineBurn3 <- pineBurn2[with(pineBurn2, order(treatment)),]
```

While I started by trying n=100, it turns out I only needed one more sample in order to get a significant result. This means 3 samples of each treatment type.

```{r}
A2 <- aov(pineBurn2$seedling_counts1~pineBurn2$treatment)
summary(A2)
```


```{r}
ggplot(pineBurn2, aes(treatment, seedling_counts1)) +
  geom_boxplot()
```

I think it is really interesting that only one more sample of each treatment type needed to be added to yeild a stasticially significant result, as well as a fairly small change in seedling counts. This paper was published by the USFS in 1950 and did not report a p-value.