---
title: "Homework 11"
author: "Kathleen Stutzman"
date: "2023-04-12"
output: html_document
---

The files I am working with are from my own research/field work assessing regeneration in Pitch Pine Barrens across the northeast. Used here are counts of small seedlings (less than 50cm) from 1 square meter frame plots. Each CSV file represents a different treatment type. 

Hidden below are `library(tidyverse)`, `library(dplyr)`, & `library(ggplot2)`

```{r Libraries, include=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

Getting started
```{r}
# Source files -------------------

source("AnalyzeData_fun.R")
source("GroupData_fun.R")

# Global variables -------------------

file_folder <- "DatasetsforLab/"
file_out <- "StatsSummary.csv"

# Body -------------------

file_names <- list.files(path=file_folder)

```


```{r GroupData_fun, eval=FALSE, include=FALSE}
#################################
# FUNCTION: group_data
# packages: dplyr, tidyverse
# purpose: to group data by site, create a number of plots for each site, and create a average of counts for each species found at each site
# input: pitch pine treatment data csv
# output: sorted dataframe with n unique plots and spec_mean calculated
# -------------------------------
group_data <- function(d=NULL) {
  if(is.null(d)){
    x_var <- runif(3)
    y_var<- sample(LETTERS, 3)
    d <- data.frame(x_var,y_var)
  }
  
  d <- d %>% 
    group_by(Site) %>% 
    mutate(nplots=n_distinct(Plot_Num)) %>% 
    group_by(Species_Code) %>% 
    mutate(spc_mean = (sum(Total_Num)/nplots))
  
  d <- d %>% 
    distinct(Site, .keep_all = TRUE)
  
  return(d)
  
}


```

```{r AnalyzeData_fun, eval=FALSE, include=FALSE}
#################################
# FUNCTION: analyze_data
# packages: none
# purpose: to calculate an ANOVA on species means and return summary statistics 
# input: sorted df using group_data
# output: p_val & f_stat
# -------------------------------
analyze_data <- function(d=NULL) {
  if(is.null(d)){
    x_var <- runif(3)
    y_var<- sample(LETTERS, 3)
    . <- data.frame(x_var,y_var)
  }
  
  d_aov <- aov(d$spc_mean~d$Latin_Name, data=d)
  . <- summary(d_aov)[[1]][1,4:5]
  
  return(.)
}
```

Create dataframe to put loop into
```{r}
stats_out <- expand_grid(ID=NA,
                         file_name=file_names,
                         F_val=NA,
                         P_val=NA,)
```

For loop for ANOVA analysis
```{r}
for (i in seq_along(file_names)){
  data <- read.table(file=paste(file_folder,file_names[i],sep=""),
                     sep = ",",
                     header=TRUE)
  
  . <- group_data(data)
  .two <- analyze_data(.)
  
  stats_out[i,3:4] <- .two
  stats_out[i,1] <- i
  
}

    write.table(cat("# Summary stats for ",
                    "batch processing of ANOVA","\n",
                    "# timestamp: ",as.character(Sys.time()),"\n",
                    "# KAS","\n",
                    "# ------------------------", "\n",
                    "\n",
                    file=file_out,
                    row.names="",
                    col.names="",
                    sep=""))
    
    write.table(x=stats_out,
              file=file_out,
              row.names=FALSE,
              col.names=TRUE,
              sep=",",
              append=TRUE)
```

For loop to create boxplots using ggplot

```{r}
myplots2 <- list()
 
for (i in seq_along(file_names)){
  data <- read.table(file=paste(file_folder,file_names[i],sep=""),
                     sep = ",",
                     header=TRUE)

  . <- group_data(data)

  z <- ggplot(data = .) +
    aes(x=Species_Code, y=spc_mean,fill=Site) +
    geom_point(size=2,
               shape=21,
               position="jitter")+
    labs(title=file_names[i],
         x="Species Codes",
         y="Average Abundance")+
    theme(legend.position = "bottom",
          legend.key.size = unit(0.5, 'cm'),
          axis.text.x = element_text(angle = 45),
          plot.title = element_text(hjust = 0.5))

  myplots2[[i]] <- z

  }

print(myplots2)

```

At first I couldn't get ggplot to work for printing multiple graphs, so I was able to use base r before figuring it out. 

```{r}
myplots <- list()

for (i in seq_along(file_names)){
  data <- read.table(file=paste(file_folder,file_names[i],sep=""),
                     sep = ",",
                     header=TRUE)
  
  . <- group_data(data)
  
  z <- boxplot(.$spc_mean~.$Species_Code, main=file_names[i])
  
  myplots[[i]] <- z

  }


```

