---
title: "Homework 7"
author: "KS"
date: "2023-03-08"
output: html_document
---
1. Start with random data

The library is open
```{r}
library(ggplot2) # for graphics
library(MASS)# for maximum likelihood estimation
library(tidyverse)
```

Read in vector data
```{r}
# quick and dirty, a truncated normal distribution to work on the solution set

z <- rnorm(n=3000,mean=0.2)
z <- data.frame(1:3000,z)
names(z) <- list("ID","myVar")
z <- z[z$myVar>0,]
str(z)
summary(z$myVar)
```

Plot a histogram
```{r}
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)
```

Add an empirical density curve
```{r}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```

Get maximum likelihood parameters for normal
```{r}
normPars <- fitdistr(z$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute
```

Plot normal probability density
```{r}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
 p1 + stat
```

Plot exponential probability density
```{r}
expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
 p1 + stat + stat2
```

Plot uniform probability density
```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
 p1 + stat + stat2 + stat3
```

Plot gamma probability density
```{r}
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
```

Plot beta probability density
```{r}
pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```




2. Now a second time, with some of my data


Read in vector data
```{r}
# loading in small seedling data for sampled sites that have been thinned and burned in NY & NH pitch pine barrens

z <- read.csv("Thinned_Burned_Data_Bio381.csv",header=TRUE)
z2 <- z[complete.cases(z$Total..),]

```


Plot a histogram
```{r}
z1 <- ggplot(data=z2, aes(x=Total.., y=after_stat(density))) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2, binwidth = 1) 
print(z1)
```

Add an empirical density curve
```{r}
z1 <-  z1 +  geom_density(linetype="dotted",size=0.75)
print(z1)
```

Get maximum likelihood parameters for normal
```{r}
znormPars <- fitdistr(z2$Total..,"normal", )
print(znormPars)
str(znormPars)
znormPars$estimate["mean"] # note structure of getting a named attribute
```

Plot normal probability density
```{r}
zmeanML <- znormPars$estimate["mean"]
zsdML <- znormPars$estimate["sd"]

zxval <- seq(0,max(z2$Total..),len=length(z2$Total..))

 zstat <- stat_function(aes(x = Total.., y = ..y..), fun = dnorm, colour="red", n = length(z2$Total..), args = list(mean = zmeanML, sd = zsdML))
 z1 + zstat
```

Plot exponential probability density
```{r}
zexpoPars <- fitdistr(z2$Total..,"exponential")
zrateML <- zexpoPars$estimate["rate"]

zstat2 <- stat_function(aes(x = Total.., y = ..y..), fun = dexp, colour="blue", n = length(z2$Total..), args = list(rate=zrateML))
 z1 + zstat + zstat2
```

Plot uniform probability density
```{r}
zstat3 <- stat_function(aes(x = Total.., y = ..y..), fun = dunif, colour="darkgreen", n = length(z2$Total..), args = list(min=min(z2$Total..), max=max(z2$Total..)))
 z1 + zstat + zstat2 + zstat3
```

Plot gamma probability density
```{r}
zgammaPars <- fitdistr(z2$Total..,"gamma")
zshapeML <- zgammaPars$estimate["shape"]
zrateML <- zgammaPars$estimate["rate"]

zstat4 <- stat_function(aes(x = Total.., y = ..y..), fun = dgamma, colour="brown", n = length(z2$Total..), args = list(shape=zshapeML, rate=zrateML))
 z1 + zstat + zstat2 + zstat3 + zstat4
```

Plot beta probability density
```{r}
pSpecial <- ggplot(data=z2, aes(x=Total../(max(Total.. + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z2$Total../max(z2$Total.. + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = Total.., y = ..y..), fun = dbeta, colour="orchid", n = length(z2$Total..), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

3. I think the exponential and gamma distributions fit best.

4. Now simulating the data
There are 107 observations, the rate is 0.269

Generating a random exponential distribution

```{r}
q <- rexp(107, r=0.269)
q <- data.frame(1:107,q)
names(q) <- list("ID","myVar")

qexpoPars <- fitdistr(q$myVar,"exponential")
qrateML <- qexpoPars$estimate["rate"]

q1 <- ggplot(data=q, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 

qstat2 <- stat_function(aes(x = myVar, y = ..y..), fun = dexp, colour="blue", n = length(q$myVar), args = list(rate=qrateML))
 q1 + qstat2

 z1 + zstat2
```

